---
title: "Union Membership Wages"
author: "Sharleen Weatherley"
date: "2023-09-05"
format: gfm
execute:
  echo: true
  warning: false
---

## Setup

First I will call the necessary libraries and read in the data, found at the [`tidytuesday` Github repository](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-09-05).

This data was curated by [unionstats.com](unionstats.com).

```{r}
#| output: false

# Library calls
library(dplyr)
library(visdat)
library(skimr)
library(ggplot2)
library(janitor)
library(stringr)

# Reading the data in
wages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-09-05/wages.csv')


# Set graph theme
theme_set(theme_minimal()) +
  theme_update(
    plot.caption = element_text(hjust = 0))

```


## Initial thoughts

From the data dictionary on the Github repo, it looks like the `wages` dataset will be the most interesting data for me, so I will start by getting an overview of what the dataset looks like.

## Exploring  

```{r}

wages |> 
  vis_dat()
```


Good, from this `vis_dat()` graph, I can see there's no missing variables. The data is nice and cleaned.

Now I will use the `skimr::skim()` function on the data.

```{r}

wages |> 
  skim()
```

Even from this `skim()` function, I can see that union wages tend to be higher than non-union wages!


I'll filter for all the workers, and create a graph to visualize this difference over the years.

```{r}


wages |> 
  filter(facet == "all wage and salary workers") |> 
  ggplot(aes(x = year)) +
  geom_point(aes(y = union_wage,
                 color = "Union wage")) +
  geom_point(aes(y = nonunion_wage,
                 color = "Non-union wage")) +
  labs(x = "Year",
       y = "Average wages ($)",
       color = "Legend",
       caption = "How wages for unionized and non-unionized workers compare over the years.") +
  scale_color_manual(values = c("#1dc58f",
                                "#d10707"))


```


## `facet` values

So what other values can the `facet` column have? I'll use `janitor`'s `tabyl()` function to check it out:

```{r}

wages |> 
  janitor::tabyl(facet)

```

I'm interested to see if the demographics of gender and race show any differences of wages over the years:

```{r}

wages |> 
  filter(str_detect(facet, "male")) |>
  mutate(facet = facet |> 
           str_remove("demographics: ") |>
           str_to_sentence()) |> 
  filter(!(facet %in% c("Male", "Female"))) |> 
  ggplot(aes(x = year)) +
  geom_point(aes(y = union_wage,
                 color = "Union wage")) +
  geom_point(aes(y = nonunion_wage,
                 color = "Non-union wage")) +
  labs(x = "Year",
       y = "Average wages ($)",
       color = "Legend",
       caption = "How wages for unionized and non-unionized workers compare over the years, split by gender and race.") +
  scale_color_manual(values = c("#1dc58f",
                                "#d10707")) +
  facet_wrap(facet ~ ., ncol = 2)

```




